---
title: "Lecture 1"
author: "Han Zhang"
date: "2/1/2021"
output: pdf_document
---
# create population
For learning purposes, we generate a pseudo population. 
The population is 100,000 units following Beta distribution.

Note that in real life, you don't observe the population. 
Here we assume we know them.
It's a common way to learn about statistics and design methods.

```{r}
population <- rbeta (100000, 1, 2)
plot (density (population))
# population mean
mean(population)
# population variance
var(population)

```

# create sample
We create one survey of 1000 sampled observations from the population.


We calculated sample mean $\bar{X}$ and $\hat{V}(X)$
Compare these estimates with population quantities. What do you find?

```{r}
sample = sample(population, size = 1000)
plot (density (sample))
# sample mean as estimator of population mean
est.mean = mean(sample)

# unbiased estimator of population variance
est.var = 1000/(1000-1) * var(sample)

```

# Normally-approximated confidence intervals

First we need to estimate standard error of the sample mean

We know it is

$\hat{V}(\bar{X}) = \frac{\hat{V}(X)}{n}$. We have calculated $\hat{V}(X)$ in the previous code chunk.

```{r}
sample_mean = mean(sample)

#standard error of the sample mean
ss = sqrt(est.var/1000)

print ("point estimate of mean")
print (sample_mean)
print ("95% normal-approximated confidence interval of mean")
c(sample_mean - 1.96 * ss, sample_mean + 1.96 * ss)

```


# Law of Large Numbers
We vary the size of survey, from 100 to 100000, with an interval of 1000.

In other words, $n$ changes from 100, 500, 1100, 1600, 2100, and so on.

As the plot shows, when $n$ increases, sample mean approaches population mean.

```{r}
sample_mean_list <- c()
sample_times <- seq(100, 100000, 500)
for (n in sample_times ){
  sample <- sample(population, size = n, replace = F)
  sample_mean_list <- c(sample_mean_list, mean(sample))
}

plot(sample_times, sample_mean_list, xlab = "number of samples", ylab = "sample mean")
abline(h = mean(population), col = "red")
```

#  Central limit theorem

This time, we let survey size to be fixed at 1000.

We explore three cases:
- case 1: we create 100 surveys, each of size 1000, from the population. Plot color is red.
- case 2: we create 1000 surveys, each of size 1000, from the population. Plot color is green.
- case 3: we create 10000 surveys, each of size 1000, from the population. Plot color is 10000.


- As $n$ increases, distribution of sample means approaches normal

```{r}

# Now we show the central limit theorem
sample_times <- c(100, 1000, 10000)
col <- c("red", "green", "blue")
i = 1
for (n in sample_times ){
  # central limit theorem talks about distribution of the sample mean
  # we cannot calculate the distribution for a single sample, so we draw sample multiple times
  sample_mean_list <- c()
  for (m in 1:1000)
  {
    sample <- sample(population, size = n, replace = TRUE)
    sample_mean_list <- c(sample_mean_list, mean(sample))
  }
  # sample_mean_list_standard <- sample_mean_list - mean(population)
  plot(density (sample_mean_list),  col = col[i], xlim = c(-0.5, 0.5), ylim  = c(0,50), xlab = "")
  abline(v = mean(population), col = "black")
  par(new = T)
  i = i  + 1
}


```

